---
title: "parameter-tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{parameter-tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hyperion)
library(hyperion.tables)

library(gt)
library(flextable)

model_dir <- system.file("extdata", "models", "onecmt", package = "hyperion.tables")
model_run <- "run003"
```


```{r}
spec <- TableSpec() |>
  set_spec_transforms(omega = "cv") |>
  set_spec_sections(
    kind == "THETA" ~ "Structural model parameters",
    kind == "OMEGA" & diagonal ~ "Interindividual variance parameters",
    kind == "OMEGA" & !diagonal ~ "Interindividual covariance parameters",
    kind == "SIGMA" ~ "Residual error",
    TRUE ~ "Other"
  ) |>
  set_spec_parameter_names(source = "display") |>
  set_spec_title(paste(model_run, "Parameters"))

mod <- read_model(file.path(model_dir, paste0(model_run, ".mod")))

info <- get_model_parameter_info(
	mod,
  system.file("lookup.toml", package = "hyperion.tables")
)

info <- info |> 
  update_param_info("SIGMA(2,2)", parameterization = "AddErr")

mod_sum <- summary(mod)

get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table()
```


## Changing display names with `set_spec_parameter_names()`

```{r}
spec <- spec |> set_spec_parameter_names(source = "display")

info <- info |>
	update_param_info("SIGMA(1,1)", display = "Proportional Error") |>
  update_param_info("SIGMA(2,2)", display = "Additive Error")

mod_sum <- summary(mod)

get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table()
```

```{r}
spec <- spec |> set_spec_parameter_names(source = "nonmem")

get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table()
```

### Adding descriptions with `add_spec_columns()`

```{r}
spec <- spec |>
  set_spec_parameter_names(source = "display") |>
  add_spec_columns("description")

get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table()
```

### Removing columns with `drop_spec_columns()`

```{r}
## Start fresh to demonstrate drop_columns
spec <- TableSpec() |>
  set_spec_transforms(omega = "cv") |>
  set_spec_sections(
    kind == "THETA" ~ "Structural model parameters",
    kind == "OMEGA" & diagonal ~ "Interindividual variance parameters",
    kind == "OMEGA" & !diagonal ~ "Interindividual covariance parameters",
    kind == "SIGMA" ~ "Residual error",
    TRUE ~ "Other"
  ) |>
  set_spec_parameter_names(source = "display") |>
  drop_spec_columns("unit")

get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table()
```

```{r}
## Drop multiple columns
spec <- spec |> drop_spec_columns("shrinkage")

get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table()
```

## Structural parameters only

```{r}
sp_spec <- TableSpec() |>
  set_spec_sections(
    kind == "THETA" ~ "Structural model parameters",
    TRUE ~ "Other"
  ) |>
  set_spec_filter(kind == "THETA") |>
  drop_spec_columns("shrinkage")

get_parameters(mod) |>
  apply_table_spec(sp_spec, info) |>
  make_parameter_table()
```

## Random Effect Parameters only
```{r}
re_spec <- TableSpec() |>
  set_spec_sections(
    kind == "OMEGA" ~ "Random Effect Parameters",
    kind == "SIGMA" ~ "Residual Error",
    TRUE ~ "Other"
  ) |>
  set_spec_filter(kind != "THETA") |>
  drop_spec_columns("unit")

get_parameters(mod) |>
  apply_table_spec(re_spec, info) |>
  make_parameter_table()
```

## Confidence Interval level
```{r}
spec <- TableSpec() |>
  set_spec_transforms(omega = "cv") |>
  set_spec_sections(
    kind == "THETA" ~ "Structural model parameters",
    kind == "OMEGA" & diagonal ~ "Interindividual variance parameters",
    kind == "OMEGA" & !diagonal ~ "Interindividual covariance parameters",
    kind == "SIGMA" ~ "Residual variance",
    TRUE ~ "Other"
  ) |>
  set_spec_ci(level = 0.7) |>
  set_spec_sigfig(3)

mod_sum <- summary(mod)
info <- get_model_parameter_info(mod)

get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table()
```

## Changing summary info shown

```{r}
spec <- TableSpec() |>
  set_spec_transforms(omega = "cv") |>
  set_spec_sections(
    kind == "THETA" ~ "Structural model parameters",
    kind == "OMEGA" & diagonal ~ "Interindividual variance parameters",
    kind == "OMEGA" & !diagonal ~ "Interindividual covariance parameters",
    kind == "SIGMA" ~ "Residual variance",
    TRUE ~ "Other"
  ) |>
  set_spec_ci(level = 0.7) |>
  set_spec_sigfig(3) |>
  set_spec_parameter_names(source = "display")

mod_sum <- summary(mod)
info <- get_model_parameter_info(mod, lookup_path = normalizePath("../inst/lookup.toml"))

get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum, show_cond_num = FALSE) |>
  make_parameter_table()
```

```{r}
get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum, show_cond_num = FALSE, show_ofv = FALSE) |>
  make_parameter_table()
```

```{r}
get_parameters(mod) |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum, show_method = FALSE) |>
  make_parameter_table()
```


## Transformation Reference

The following table shows how CV, RSE, and CI are computed for each transform and parameter type combination.

```{r transformation-reference, echo=FALSE}
# Default formulas
default_rse <- ""
default_ci <- ""
na_val <- "N/A"

# Build the reference table
transform_ref <- data.frame(
  transform = rep(
    c("Identity", "LogNormal", "Logit", "Proportional", "AddErr", "LogAddErr"),
    each = 3
  ),
  parameter = rep(c("Theta", "Omega", "Sigma"), 6),
  cv = c(
    # Identity: all N/A
    na_val, na_val, na_val,
    # LogNormal: Theta N/A, Omega/Sigma have formula
    na_val,
    "$\\sqrt{\\exp(\\text{Est}) - 1}$",
    "$\\sqrt{\\exp(\\text{Est}) - 1}$",
    # Logit: all N/A
    na_val, na_val, na_val,
    # Proportional: Theta N/A, Omega/Sigma have formula
    na_val,
    "$\\sqrt{\\text{Est}}$",
    "$\\sqrt{\\text{Est}}$",
    # AddErr: all N/A
    na_val, na_val, na_val,
    # LogAddErr: Theta has formula, Omega N/A, Sigma has formula
    "$\\sqrt{\\exp(\\text{Est}^2) - 1}$",
    na_val,
    "$\\sqrt{\\exp(\\text{Est}) - 1}$"
  ),
  rse = c(
    # Identity: all default
    default_rse, default_rse, default_rse,
    # LogNormal: Theta special, Omega/Sigma default
    "$\\sqrt{\\exp(\\text{SE}^2) - 1}$",
    default_rse, default_rse,
    # Logit: Theta special, Omega/Sigma default
    "$(1 - \\text{BT}) \\cdot \\text{SE}$",
    default_rse, default_rse,
    # Proportional: all default
    default_rse, default_rse, default_rse,
    # AddErr: all default
    default_rse, default_rse, default_rse,
    # LogAddErr: all default
    default_rse, default_rse, default_rse
  ),
  ci = c(
    # Identity
    default_ci, default_ci, default_ci,
    # LogNormal
    "$\\exp(\\text{Est} \\pm z \\cdot \\text{SE})$",
    "$\\exp(\\text{Est} \\pm z \\cdot \\text{SE})$",
    "$\\exp(\\text{Est} \\pm z \\cdot \\text{SE})$",
    # Logit
    "$\\frac{1}{1 + \\exp(-(\\text{Est} \\pm z \\cdot \\text{SE}))}$",
    "$\\frac{1}{1 + \\exp(-(\\text{Est} \\pm z \\cdot \\text{SE}))}$",
    "$\\frac{1}{1 + \\exp(-(\\text{Est} \\pm z \\cdot \\text{SE}))}$",
    # Proportional
    default_ci, default_ci, default_ci,
    # AddErr
    default_ci, default_ci, default_ci,
    # LogAddErr
    default_ci, default_ci, default_ci
  ),
  stringsAsFactors = FALSE
)

transform_ref |>
  gt::gt(groupname_col = "transform") |>
  gt::cols_label(
    parameter = "Parameter",
    cv = "CV",
    rse = "RSE",
    ci = "CI"
  ) |>
  gt::fmt_markdown() |>
  gt::tab_header(title = "Transformation Formulas") |>
  gt::tab_footnote("Est = THETA(x)/OMEGA(i,j)/SIGMA(i,j) reported in the .ext file") |>
  gt::tab_footnote("BT = back-transformed estimate = 1/(1 + exp(-Est))") |>
  gt::tab_footnote("SE = Standard Error, z = z-score for CI level") |>
  gt::tab_footnote("CV and RSE formulas are multiplied by 100 to express as percentages") |>
  gt::tab_footnote(
    "SE/|Est| unless otherwise noted",
    locations = gt::cells_column_labels(columns = "rse")
  ) |>
  gt::tab_footnote(
    "Est ± z·SE unless otherwise noted",
    locations = gt::cells_column_labels(columns = "ci")
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = list(
      gt::cells_column_labels(dplyr::everything()),
      gt::cells_title(groups = "title"),
      gt::cells_row_groups()
    )
  ) |>
  gt::tab_options(
    row_group.as_column = TRUE # This option makes the group appear as a column
  )
```

## Model Lineage

```{r}
example_tree <- get_model_lineage(model_dir)

example_tree
```

```{r}
spec <- SummarySpec() |>
  set_spec_pvalue(scientific = FALSE)

example_tree |>
  apply_summary_spec(spec) |>
  make_summary_table()
```
