---
title: "table-rendering"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{table-rendering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hyperion)
library(hyperion.tables)

data_dir <- system.file("extdata", package = "hyperion")
model_dir <- file.path(data_dir, "models", "onecmt")
model_run <- "run003"
```

# Rendering and extending tables

`hyperion.tables` renders tables through a renderer-agnostic intermediate
called `HyperionTable`. The table data is produced once from a `TableSpec`,
then any renderer (gt/flextable/custom) can format and display it consistently.

Renderers call `apply_formatting()` internally, so the same rendering rules
apply across outputs.

# Pipeline overview

Here’s what happens when you run:

```
params |>
  apply_table_spec(spec, info) |>
  add_summary_info(mod_sum) |>
  make_parameter_table()
```

1. **`apply_table_spec()`** enriches the raw parameter data with transforms,
   CI bounds, section labels, and attaches `table_spec`. It does *not* create
   display strings.
2. **`add_summary_info()`** attaches model summary metadata for footnotes.
3. **`make_parameter_table()`** builds a `HyperionTable`, which captures
   labels, hide/show rules, CI merge specs, footnotes, and other display intent.
4. **`render_to_gt()` / `render_to_flextable()`** call `apply_formatting()` to:
   - format numeric columns
   - build the variability display column (if requested)
   - merge CI bounds (if enabled)
   - apply missing-value text

This keeps formatting centralized and consistent across renderers while
keeping the underlying data pipeline clean and composable.

# Build a `HyperionTable` and render it

Render rules live on `TableSpec`:

- `ci` controls CI merging and missing-value display.
- `n_sigfig` controls numeric precision.
- `missing_text` + `missing_apply_to` control `NA` substitution.

```{r}
spec <- TableSpec(
  sections = section_rules(
    kind == "THETA" ~ "Structural model parameters",
    kind == "OMEGA" & diagonal ~ "Interindividual variance parameters",
    kind == "OMEGA" & !diagonal ~ "Interindividual covariance parameters",
    kind == "SIGMA" ~ "Residual error",
    TRUE ~ "Other"
  ),
  title = paste(model_run, "Parameters")
)

run003 <- read_model(file.path(model_dir, paste0(model_run, ".mod")))

hyperion_table <- get_parameters(run003) |>
  apply_table_spec(spec, get_model_parameter_info(run003)) |>
  make_parameter_table(output = "data")

hyperion_table
```

`hyperion.tables` includes two built-in renderers:

```{r}
hyperion.tables::render_to_gt(hyperion_table)
```

```{r}
hyperion.tables::render_to_flextable(hyperion_table)
```

# Update render rules and re-render

To change formatting, rebuild the `HyperionTable` with a new `TableSpec`
and re-render it. You can also inspect the formatted data directly.

```{r}
spec <- TableSpec(
  sections = section_rules(
    kind == "THETA" ~ "Structural model parameters",
    kind == "OMEGA" & diagonal ~ "Interindividual variance parameters",
    kind == "OMEGA" & !diagonal ~ "Interindividual covariance parameters",
    kind == "SIGMA" ~ "Residual error",
    TRUE ~ "Other"
  ),
  title = paste(model_run, "Parameters"),
  n_sigfig = 4,
  ci = CIOptions(
    merge = TRUE,
    pattern = "(%s; %s)",
    missing_text = "NA"
  ),
  missing_text = "NA",
  missing_apply_to = "numeric"
)

hyperion_table <- get_parameters(run003) |>
  apply_table_spec(spec, get_model_parameter_info(run003)) |>
  make_parameter_table(output = "data")

formatted <- apply_formatting(hyperion_table)
names(formatted)
```

```{r}
hyperion.tables::render_to_gt(hyperion_table)
```


```{r}
head(formatted)
```

# Extending with a custom renderer

Custom renderers should accept a `HyperionTable`, call `apply_formatting()`,
and then build output using your chosen table package.

```{r}
render_custom <- function(table) {
  data <- apply_formatting(table)
  # Replace with your rendering package of choice
	gt::gt(data, groupname_col = table@groupname_col)
}

render_custom(hyperion_table)
```

# What if you skip `apply_formatting()`?

If you render directly from the raw `HyperionTable@data`, you will see
unformatted numeric values and separate CI columns (if present). This is
useful for debugging, but not recommended for final output.

```{r}
raw_data <- hyperion_table@data
gt::gt(raw_data, groupname_col = hyperion_table@groupname_col)
```

# Notes

- The render rules are **shared** across renderers to keep tables consistent.
- `make_parameter_table(output = "data")` returns the `HyperionTable`.
- If you need tighter control, use `apply_formatting()` and pass the returned
  data frame to your custom renderer.
